---
title: "The Normal CDF and pnorm"
output: html_notebook
---
Key points
The normal distribution has a mathematically defined CDF which can be computed in R with the function pnorm().
pnorm(a, avg, s) gives the value of the cumulative distribution function F(a) for the normal distribution defined by average avg and standard deviation s.
We say that a random quantity is normally distributed with average avg and standard deviation s if the approximation pnorm(a, avg, s) holds for all values of a.
If we are willing to use the normal approximation for height, we can estimate the distribution simply from the mean and standard deviation of our values.
If we treat the height data as discrete rather than categorical, we see that the data are not very useful because integer values are more common than expected due to rounding. This is called discretization.
With rounded data, the normal approximation is particularly useful when computing probabilities of intervals of length 1 that include exactly one integer.

## Code: Using pnorm to calculate probabilities
Given male heights x:

```{r}
library(tidyverse)
library(dslabs)
data(heights)
x <- heights %>% filter(sex=="Male") %>% pull(height)
```

## We can estimate the probability that a male is taller than 70.5 inches with:
```{r}
1 - pnorm(70.5, mean(x), sd(x))
```

## Code: Discretization and the normal approximation
```{r}
# plot distribution of exact heights in data
plot(prop.table(table(x)), xlab = "a = Height in inches", ylab = "Pr(x = a)")

# probabilities in actual data over length 1 ranges containing an integer
mean(x <= 68.5) - mean(x <= 67.5)
mean(x <= 69.5) - mean(x <= 68.5)
mean(x <= 70.5) - mean(x <= 69.5)

# probabilities in normal approximation match well
pnorm(68.5, mean(x), sd(x)) - pnorm(67.5, mean(x), sd(x))
pnorm(69.5, mean(x), sd(x)) - pnorm(68.5, mean(x), sd(x))
pnorm(70.5, mean(x), sd(x)) - pnorm(69.5, mean(x), sd(x))

# probabilities in actual data over other ranges don't match normal approx as well
mean(x <= 70.9) - mean(x <= 70.1)
pnorm(70.9, mean(x), sd(x)) - pnorm(70.1, mean(x), sd(x))
```






