---
title: "Assessment: Introduction to Discrete Probability"
output: html_notebook
---

### Exercise 1. Probability of cyan - generalized
In the edX exercises for this section, we calculated some probabilities by hand. Now we'll calculate those probabilities using R.

One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball will be cyan?

Instructions
100 XP
Define a variable p as the probability of choosing a cyan ball from the box.
Print the value of p.


```{r}
cyan <- 3
magenta <- 5
yellow <- 7

# Assign a variable `p` as the probability of choosing a cyan ball from the box

p<-cyan/(cyan+magenta+yellow)

# Print the variable `p` to the console
p
```


### Exercise 2. Probability of not cyan - generalized
We defined the variable p as the probability of choosing a cyan ball from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball you draw from the box will NOT be cyan?

Instructions
100 XP
Using the probability of choosing a cyan ball, p, calculate the probability of choosing any other ball.


```{r}
Exercise 2. Probability of not cyan - generalized
We defined the variable p as the probability of choosing a cyan ball from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball you draw from the box will NOT be cyan?

Instructions
100 XP
Using the probability of choosing a cyan ball, p, calculate the probability of choosing any other ball.

```


### Exercise 3. Sampling without replacement - generalized
Instead of taking just one draw, consider taking two draws. You take the second draw without returning the first draw to the box. We call this sampling without replacement.

What is the probability that the first draw is cyan and that the second draw is not cyan?

Instructions
100 XP
Calculate the conditional probability p_2 of choosing a ball that is not cyan after one cyan ball has been removed from the box.
Calculate the joint probability of both choosing a cyan ball on the first draw and a ball that is not cyan on the second draw using p_1 and p_2.


```{r}
cyan <- 3
magenta <- 5
yellow <- 7

# The variable `p_1` is the probability of choosing a cyan ball from the box on the first draw.
p_1 <- cyan / (cyan + magenta + yellow)

# Assign a variable `p_2` as the probability of not choosing a cyan ball on the second draw without replacement.

p_2 <- (magenta + yellow ) / (cyan + magenta + yellow -1)

# Calculate the probability that the first draw is cyan and the second draw is not cyan using `p_1` and `p_2`.
p_1*p_2
```


### Exercise 4. Sampling with replacement - generalized
Now repeat the experiment, but this time, after taking the first draw and recording the color, return it back to the box and shake the box. We call this sampling with replacement.

What is the probability that the first draw is cyan and that the second draw is not cyan?

Instructions
100 XP
Calculate the probability p_2 of choosing a ball that is not cyan on the second draw, with replacement.
Next, use p_1 and p_2 to calculate the probability of choosing a cyan ball on the first draw and a ball that is not cyan on the second draw (after replacing the first ball).


```{r}
cyan <- 3
magenta <- 5
yellow <- 7

# The variable 'p_1' is the probability of choosing a cyan ball from the box on the first draw.
p_1 <- cyan / (cyan + magenta + yellow)

# Assign a variable 'p_2' as the probability of not choosing a cyan ball on the second draw with replacement.

p_2 <- 1-p_1

# Calculate the probability that the first draw is cyan and the second draw is not cyan using `p_1` and `p_2`.

p_1*p_2
```
