---
title: "Combinations and Permutations"
output: html_notebook
---

Key points
paste() joins two strings and inserts a space in between.
expand.grid() gives the combinations of 2 vectors or lists.
permutations(n,r) from the gtools package lists the different ways that r items can be selected from a set of n options when order matters.
combinations(n,r) from the gtools package lists the different ways that r items can be selected from a set of n options when order does not matter.
Code: Introducing paste() and expand.grid()
# joining strings with paste
```{r}
number <- "Three"
suit <- "Hearts"
paste(number, suit)
```
# joining vectors element-wise with paste
```{r}
paste(letters[1:5], as.character(1:5))
```
# generating combinations of 2 vectors with expand.grid
```{r}
expand.grid(pants = c("blue", "black"), shirt = c("white", "grey", "plaid"))
Code: Generating a deck of cards
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
numbers <- c("Ace", "Deuce", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten", "Jack", "Queen", "King")
deck <- expand.grid(number = numbers, suit = suits)
deck <- paste(deck$number, deck$suit)
```
# probability of drawing a king
```{r}
kings <- paste("King", suits)
mean(deck %in% kings)
```
Code: Permutations and combinations
Correction: The code shown does not generate all 7 digit phone numbers because phone numbers can have repeated digits. It generates all possible 7 digit numbers without repeats.
```{r}
library(gtools)
permutations(5,2)    # ways to choose 2 numbers in order from 1:5
all_phone_numbers <- permutations(10, 7, v = 0:9)
n <- nrow(all_phone_numbers)
index <- sample(n, 5)
all_phone_numbers[index,]

permutations(3,2)    # order matters
combinations(3,2)    # order does not matter

```
Code: Probability of drawing a second king given that one king is drawn
```{r}
hands <- permutations(52,2, v = deck)
first_card <- hands[,1]
second_card <- hands[,2]
sum(first_card %in% kings)

sum(first_card %in% kings & second_card %in% kings) / sum(first_card %in% kings)
```
Code: Probability of a natural 21 in blackjack
```{r}
aces <- paste("Ace", suits)
facecard <- c("King", "Queen", "Jack", "Ten")
facecard <- expand.grid(number = facecard, suit = suits)
facecard <- paste(facecard$number, facecard$suit)

hands <- combinations(52, 2, v=deck) # all possible hands

# probability of a natural 21 given that the ace is listed first in `combinations`
```{r}
mean(hands[,1] %in% aces & hands[,2] %in% facecard)
```
# probability of a natural 21 checking for both ace first and ace second
```{r}
mean((hands[,1] %in% aces & hands[,2] %in% facecard)|(hands[,2] %in% aces & hands[,1] %in% facecard))
```
Code: Monte Carlo simulation of natural 21 in blackjack
Note that your exact values will differ because the process is random and the seed is not set.

# code for one hand of blackjack
```{r}
hand <- sample(deck, 2)
hand
```
# code for B=10,000 hands of blackjack
```{r}
B <- 10000
results <- replicate(B, {
    hand <- sample(deck, 2)
    (hand[1] %in% aces & hand[2] %in% facecard) | (hand[2] %in% aces & hand[1] %in% facecard)
})
mean(results)
```

```{r}

```
