---
title: "Assessment: Quantiles, percentiles, and boxplots"
output: html_notebook
---

## Exercise 1. Vector lengths
When analyzing data it's often important to know the number of measurements you have for each category.

Instructions
Define a variable male that contains the male heights.
Define a variable female that contains the female heights.
Report the length of each variable.

```{r}
library(dslabs)
data(heights)
male <- heights$height[heights$sex=="Male"]
female <- heights$height[heights$sex=="Female"]
length(male)
length(female)
```
## Exercise 2. Percentiles
Suppose we can't make a plot and want to compare the distributions side by side. If the number of data points is large, listing all the numbers is inpractical. A more practical approach is to look at the percentiles. We can obtain percentiles using the quantile function like this

library(dslabs)
data(heights)
quantile(heights$height, seq(.01, 0.99, 0.01))

Instructions
Create two five row vectors showing the 10th, 30th, 50th, 70th, and 90th percentiles for the heights of each sex called these vectors female_percentiles and male_percentiles.
Then create a data frame called df with these two vectors as columns. The column names should be female and male and should appear in that order. As an example consider that if you want a data frame to have column names names and grades, in that order, you do it like this:
df <- data.frame(names = c("Jose", "Mary"), grades = c("B", "A"))
Take a look at the df by printing it. This will provide some information on how male and female heights differ.

```{r}
library(dslabs)
data(heights)
male <- heights$height[heights$sex=="Male"]
female <- heights$height[heights$sex=="Female"]
p <- c(0.1,0.3,0.5,0.70,0.90)
female_percentiles <- quantile(female, p)
male_percentiles <- quantile(male, p)
df <- data.frame(female = female_percentiles,male =male_percentiles)
df
```

## Exercise 3. Interpreting Boxplots - 1
Study the boxplots summarizing the distributions of populations sizes by country.

Which continent has the country with the largest population size?

Instructions


Africa
Americas
Asia
Europe
Oceania

```{r}
```

```{r}
```

```{r}
```

```{r}
```

